### 2020-10-18 Kommentarer.Rmd
Vi har fjernet Appendiksen, ellers er oppgaven fin og god! Vi har også lagt inn kommentarer!

---
title: "Kommentarer"
output: html_document
---

Hei!